---
title: "i.Exercise2.4.8"
author: "Kaarthik, Sahil and Vidhi"
date: "5/24/2020"
output: 
  pdf_document:
    latex_engine: xelatex

header~include:
  \usepackage{float}
  \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This exercise relates to the **College** data set, which can be found in
the file **College.csv**. It contains a number of variables for 777 different
universities and colleges in the US.

(a) Use the `read.csv()` function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.

```{r parta}
getwd()
college <- read.csv("College.csv")
```

(b) Look at the data using the `fix()` function. You should notice
that the first column is just the name of each university. We don’t
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:

```{r partb1}
#fix(college)
rownames ( college )= college [,1]
college = college [,-1]
##fix ( college )
```

```{r partb2}
library(knitr)
library(kableExtra)
kable(head(college), format = "latex", booktabs=T) %>%
  kable_styling(latex_options="scale_down")
# head(college)
# library()
```

(c) i. Use the `summary()` function to produce a numerical summary
of the variables in the data set.

```{r partc1}
summary(college)
```

(c) ii. Use the `pairs()` function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].

```{r partc2}
pairs(college[,1:10])
```

(c) iii. Use the `plot()` function to produce side-by-side boxplots of
Outstate versus Private

```{r partc3}
attach(college)
plot(Outstate~Private, data=college, xlab = "Private", ylab = "Outstate")
```

(c) iv. Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10 % of their high school
classes exceeds 50 %.

```{r partc4}
Elite = rep("No", nrow(college))
Elite[Top10perc>50] = "Yes"
#Elite
Elite = as.factor(Elite)
college = data.frame(college, Elite)
summary(college)
plot(Outstate~Elite, xlab = "Elite", ylab = "Outstate")
```

(c) v. Use the `hist()` function to produce some histograms with
differing numbers of bins for a few of the quantitative variables. You may find the command `par(mfrow=c(2,2))` useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.

```{r partc5}
par(mfrow=c(2,2))
hist(Apps, breaks=50, xlim=c(0,25000), main="Apps")
hist(Enroll, breaks = 25, main = "Enroll")
hist(Expend, breaks = 25, main = "Expend")
hist(Outstate, breaks = 25, main = "Outstate")
```

(c) vi. Continue exploring the data, and provide a brief summary
of what you discover.

```{r partc6}
table(Private, Elite)
library(gmodels)
CrossTable(Private, Elite)
```

According to the scatter plot from the **College** dataset, it is really hard to find the correlation in the small screen as the laptop. Although we can zoom in and find the relationship between the variables or plot the single graph with providing two variables as x and y using the `plot` function. From the variable Elite that we created that only hold the top 10 percent more than 50 as in yes. From the table shown above it generates the confusion matrix between the Elite and Private University.

```{r plot}
plot(college$F.Undergrad, college$Enroll)
```