---
title: "ix. Exercise 10.7.10"
author: "Kaarthik Sundaramoorthy, Sahil Shah and Vidhi Shah"
date: "7/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this problem, you will generate simulated data, and then perform
PCA and K-means clustering on the data.

(a) Generate a simulated data set with 20 observations in each of
three classes (i.e. 60 observations total), and 50 variables.

Hint: There are a number of functions in `R` that you can use to
generate data. One example is the `rnorm()` function; `runif()` is
another option. Be sure to add a mean shift to the observations
in each class so that there are three distinct classes.

```{r datagen}
set.seed(3)
x = matrix(rnorm(20*3*50, mean = 0, sd = 0.01), ncol = 50)
x[1:20, 2] = 1
x[21:40, 1] = 2
x[21:40, 2] = 2
x[41:60, 1] = 1
true.labels <- c(rep(1,20), rep(2,20), rep(3,20))
summary(x)
```

(b) Perform PCA on the 60 observations and plot the first two principal component score vectors. Use a different color to indicate the observations in each of the three classes. If the three classes appear separated in this plot, then continue on to part (c). If
not, then return to part (a) and modify the simulation so that there is greater separation between the three classes. Do not continue to part (c) until the three classes show at least some
separation in the first two principal component score vectors.

```{r partb}
pca.out = prcomp(x)
summary(pca.out)
pca.out$x[,1:2]
plot(pca.out$x[,1:2], col = 2:4, xlab = 'Z1', ylab = 'Z2', pch = 20)
```

(c) Perform K-means clustering of the observations with K = 3. How well do the clusters that you obtained in K-means clustering compare to the true class labels?

Hint: You can use the `table()` function in R to compare the true
class labels to the class labels obtained by clustering. Be careful
how you interpret the results: K-means clustering will arbitrarily
number the clusters, so you cannot simply check whether the true
class labels and clustering labels are the same.

```{r partc}
library(cluster)
library(fpc)
kmeans.out.3 = kmeans(x, 3, nstart = 20)
table(true.labels, kmeans.out.3$cluster)
plotcluster(x, kmeans.out.3$cluster)
clusplot(x, kmeans.out.3$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

There are three clusters defined in the dataset with 20 observation each. They are perfectly clustered in three classes from the `kmeans()` clustering. In the plot you can see there are three clusters numbered 1,2 and 3. 

(d) Perform K-means clustering with K = 2. Describe your results.

```{r partd}
kmeans.out.2 = kmeans(x, 2, nstart = 20)
table(true.labels, kmeans.out.2$cluster)
plotcluster(x, kmeans.out.2$cluster)
clusplot(x, kmeans.out.2$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

While we define **k = 2**. The two clusters point merge into one cluster. As you can see in the plot, the cluster 1 and 2 from previous plot merged into 1 cluster, Leaving **20** observations in one and **40** observations in the other cluster.

(e) Now perform K-means clustering with K = 4, and describe your
results.

```{r parte}
kmeans.out.4 = kmeans(x, 4, nstart = 20)
table(true.labels, kmeans.out.4$cluster)
clusplot(x, kmeans.out.4$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

The defined dataset with 3 classes are now classified into 4 clusters. With __cluster-2__ and __cluster-3__ perfectly clustered with __20__ observartion each and other two contains the distributed observations of third one, with __9__ in __cluster-4__ and __11__ in __cluster-1__.

(f) Now perform K-means clustering with K = 3 on the first two
principal component score vectors, rather than on the raw data.
That is, perform K-means clustering on the 60 × 2 matrix of
which the first column is the first principal component score
vector, and the second column is the second principal component
score vector. Comment on the results.

```{r partf}
kmeans.out.31 = kmeans(pca.out$x[,1:2], 3, nstart=20)
table(true.labels, kmeans.out.31$cluster)
clusplot(x, kmeans.out.31$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

The results show the correctly classified instances into three clusters, with **20** observation each. Perfectly clustered again.

(g) Using the `scale()` function, perform K-means clustering with
K = 3 on the data after scaling each variable to have standard
deviation one. How do these results compare to those obtained
in (b)? Explain.

```{r partg}
kmeans.scale<-kmeans(scale(x),3,nstart=20)
table(true.labels,kmeans.scale$cluster)
clusplot(x, kmeans.scale$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

Here in the truth table we can see that the observations are not perfectly clustered. The results are not good compared to the unscaled clustering(b). The `scale()` function affects the **distances** between the observations. 