---
title: "Exercise 6.8.10"
author: "Kaarthik Sundaramoorthy, Sahil Shah and Vidhi Shah"
date: "6/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We have seen that as the number of features used in a model increases, the training error will necessarily decrease, but the test error may not. We will now explore this in a simulated data set.

(a) Generate a data set with p = 20 features,n = 1,000 observations, and an associated quantitative response vector generated according to the model Y = X $\beta$ + \(\epsilon\), where $\beta$ has some elements that are exactly equal to zero.

```{r parta}
library(leaps)
set.seed(1)
p = 20
n = 1000
x = matrix(rnorm(n * p), n, p)
B = rnorm(p)
B[3] = 0
B[4] = 0
B[9] = 0
B[19] = 0
B[10] = 0
eps = rnorm(p)
y = x %*% B + eps
```

(b) Split your data set into a training set containing 100 observations and a test set containing 900 observations.

```{r partb}
train = sample(seq(1000), 100, replace = FALSE)
y.train = y[train, ]
y.test = y[-train, ]
x.train = x[train, ]
x.test = x[-train, ]
```

(c) Perform best subset selection on the training set, and plot the training set MSE associated with the best model of each size.

```{r partc}
library(leaps)
regfit.full = regsubsets(y ~ ., data = data.frame(x = x.train, y = y.train), 
    nvmax = p)
val.errors = rep(NA, p)
x_cols = colnames(x, do.NULL = FALSE, prefix = "x.")
for (i in 1:p) {
    coefi = coef(regfit.full, id = i)
    pred = as.matrix(x.train[, x_cols %in% names(coefi)]) %*% coefi[names(coefi) %in% 
        x_cols]
    val.errors[i] = mean((y.train - pred)^2)
}
plot(val.errors, ylab = "Training MSE", pch = 19, type = "b", col = "blue")
```

(d) Plot the test set MSE associated with the best model of each
size. 

```{r partd}
val.errors = rep(NA, p)
for (i in 1:p) {
    coefi = coef(regfit.full, id = i)
    pred = as.matrix(x.test[, x_cols %in% names(coefi)]) %*% coefi[names(coefi) %in% 
        x_cols]
    val.errors[i] = mean((y.test - pred)^2)
}
plot(val.errors, ylab = "Test MSE", pch = 19, type = "b", col = "red")
```

(e) For which model size does the test set MSE take on its minimum
value? Comment on your results. If it takes on its minimum value
for a model containing only an intercept or a model containing
all of the features, then play around with the way that you are
generating the data in (a) until you come up with a scenario in
which the test set MSE is minimized for an intermediate model
size.

```{r parte}
which.min(val.errors)
```

16 parameter model has the smallest test MSE.

(f) How does the model at which the test set MSE is minimized
compare to the true model used to generate the data? Comment
on the coefficient values.

```{r partf}
coef(regfit.full, id = 18)
```

(g) 

```{r partg}
val.errors = rep(NA, p)
A = rep(NA, p)
B = rep(NA, p)
for (i in 1:p) {
    coefi = coef(regfit.full, id = i)
    A[i] = length(coefi) - 1
    B[i] = sqrt(sum((b[x_cols %in% names(coefi)] - coefi[names(coefi) %in% x_cols])^2) + 
        sum(b[!(x_cols %in% names(coefi))])^2)
}
plot(x = A, y = B, xlab = "number of coefficients", ylab = "error between estimated and true coefficients", col = "purple")
which.min(B)
```